# Topic 01

Heteroskedasticity and autocorrelation in systems of equations.

## Motivation

Analyse the effect of heteroskdeasticty in systems of equations on SOLS estimator.

## Probelm

Simulate a bivariate heteroskedasticity model for $T=2000$ observations

\begin{align*}
B y_t + A x_t &= u_t \\
u_t &\sim N\left(0, V_t = S_t S_t^{'}\right) \\
S_t &= C + D w_t
\end{align*}

with,

\begin{align*}
B &= \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}, \\
A &= \begin{bmatrix} -0.4 & 0 \\ 0 & 0.5 \end{bmatrix}, \\
C &= \begin{bmatrix} 1 & 0 \\ 0.5 & 2 \end{bmatrix}, \\
D &= \begin{bmatrix} 0.5 & 0 \\ 0.2 & 0.2 \end{bmatrix},
\end{align*}

and with,

\begin{align*}
x_{1,t} &\sim iid \,\,\, U\left[0, 10\right], \\
x_{2,t} &\sim iid \,\,\, N\left(0, 9\right), \\
w_{1,t} &\sim iid \,\,\, U\left[0, 1\right],
\end{align*}

## Analysis

Simulation of the data (*without* contemporaneous relationships in $y$)

```{r}

t <- 2000

beta1  <- 0 # 0.6
alpha1 <- 0.4
beta2  <- 0 # 0.2
alpha2 <- -0.5

c11 <- 1.0
c21 <- 0.5
c22 <- 2.0

d11  <- 0.5
d21  <- 0.2
d22  <- 0.2

b <-  matrix(c(1, -beta2,
               -beta1, 1), nrow=2, byrow=T)
a <- matrix(c(-alpha1, 0,
              0, -alpha2), nrow=2, byrow=T)
c <-  matrix(c(c11,  0,
               c21, c22), nrow=2, byrow=T)
d <-  matrix(c(d11,  0,
               d21,  d22), nrow=2, byrow=T)
# Exogenous variables
x <- cbind(10*runif(t), 3*rnorm(t))
w <- runif(t)

# Disturbances
zeros <- array(0, c(t,2))
u <- zeros
for (i in seq(t)) {
  l     <- c + d * w[i]
  u[i,] <- rnorm(2) %*% t(l)
}

# Simulate the reduced form
y <- zeros
for (i in seq(t)) {
  y[i,] <- -x[i,] %*% a %*% solve(b) + u[i,] %*% solve(b)
}

```

Estimation of $A$ using OLS (*without* contemporaneous relationships in $y$)

```{r}

# OLS fitting
lm.res <- lm(y ~ x - 1)

# compare estimates
t(lm.res$coefficients)

b.a <- solve(b) %*% -a
b.a

```

Estimation of $V$ (*without* contemporaneous relationships in $y$)

```{r}

Sig.u <- 1/nrow(lm.res$residuals) * t(lm.res$residuals) %*% lm.res$residuals
Sig.u

# reconstruct (average) Sig
l <- c + d * 0.5
V <- l %*% t(l)
V

```

<hr>

Simulation of the data (*with* contemporaneous relationships in $y$)

```{r}

t <- 2000

beta1  <- 0.6
alpha1 <- 0.4
beta2  <- 0.2
alpha2 <- -0.5

c11 <- 1.0
c21 <- 0.5
c22 <- 2.0

d11  <- 0.5
d21  <- 0.2
d22  <- 0.2

b <-  matrix(c(1, -beta2,
               -beta1, 1), nrow=2, byrow=T)
a <- matrix(c(-alpha1, 0,
              0, -alpha2), nrow=2, byrow=T)
c <-  matrix(c(c11,  0,
               c21, c22), nrow=2, byrow=T)
d <-  matrix(c(d11,  0,
               d21,  d22), nrow=2, byrow=T)
# Exogenous variables
x <- cbind(10*runif(t), 3*rnorm(t))
w <- runif(t)

# Disturbances
zeros <- array(0, c(t,2))
u <- zeros
for (i in seq(t)) {
  l     <- c + d * w[i]
  u[i,] <- rnorm(2) %*% t(l)
}

# Simulate the reduced form
y <- zeros
for (i in seq(t)) {
  y[i,] <- -x[i,] %*% a %*% solve(b) + u[i,] %*% solve(b)
}

```

Estimation of the parameters using OLS (*with* contemporaneous relationships in $y$)

```{r}

# OLS fitting
lm.res <- lm(y ~ x - 1)

# compare estimates
t(lm.res$coefficients)

b.a <- solve(b) %*% -a
b.a

```

Estimation of $V$ (*with* contemporaneous relationships in $y$)

```{r}

Sig.u <- 1/nrow(lm.res$residuals) * t(lm.res$residuals) %*% lm.res$residuals
Sig.u

# reconstruct (average) Sig
l <- c + d * 0.5
V <- l %*% t(l)
V

```

Cholesky decomposition of $V$

```{r}

S <- chol(Sig.u)
S

t(S) %*% S
Sig.u

```



