
# Topic 02

```{r,include=FALSE}

library(knitr)
knitr::opts_chunk$set(
  include = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = TRUE
  # fig.height = 8,
  # fig.width = 10,
  # fig.align = 'center'
  )
# options(scipen=1)
# options(width=9999)
# # See: https://stackoverflow.com/questions/36230790/missing-horizontal-scroll-bar-in-r-markdown-html-code-chunks-and-output

```

## Statistical Model

Statistical model for each observations $i$ (same number of $k$ regressors across $m$ equations),

\begin{align*}
\underset{\left(m \times 1\right)}{y_{i}} &= \underset{\left(m\times mk\right)}{\overline{X}_{i}} \underset{\left(mk \times 1\right)}{\beta} + \underset{\left(m \times 1\right)}{e_{i}}, \\

\begin{bmatrix} 
y_{1i} \\
y_{2i} \\
\vdots \\
y_{mi}
\end{bmatrix}
&=
\begin{bmatrix} 
x_{1i}^{'} & 0 & \cdots & 0 \\
0 & x_{2i}^{'} & \cdots & 0 \\
\vdots & \vdots & \ddots & 0 \\
0 & 0& \cdots & x_{mi}^{'}
\end{bmatrix}
+
\begin{bmatrix} 
\beta_{1} \\
\beta_{2} \\
\vdots \\
\beta_{m}
\end{bmatrix}
+
\begin{bmatrix} 
u_{1i} \\
u_{2i} \\
\vdots \\
u_{mi}
\end{bmatrix},
\end{align*}

with,

* $y_{ji}$ and $e_{ji}$ are scalars for $j=1,...,m$.
* $x_{ji}$ are $\left(k \times 1\right)$ matrix for $j=1,...,m$.
* $\beta$ are $\left(k \times 1\right)$ matrix for $j=1,...,m$.

With the same $k$ regressors across $m$ equations this simplifies to,

\begin{align*}
\underset{\left(m \times 1\right)}{y_{i}} &= \underset{\left(m\times mk\right)}{\left(\underset{\left(m \times m\right)}{I_{m}} \otimes \underset{\left(1\times k\right)}{x_{i}^{'}}\right)} \underset{\left(mk \times 1\right)}{\beta} + \underset{\left(m \times 1\right)}{e_{i}},
\end{align*}

thus,

\begin{align*}
\overline{X}_{i} =\underset{\left(m \times mk\right)}{\underset{\left(m \times m\right)}{I_{m}} \otimes \underset{\left(1 \times k\right)}{x_{i}^{'}}}
\end{align*}

Statistical model in matrix notation across observations $i$ (same $k$ regressors across $m$ equations),

\begin{align*}
\underset{\left(n \times m\right)}{Y} &= \underset{\left(n\times k\right)}{X} \underset{\left(k \times m\right)}{B} + \underset{\left(n \times k\right)}{E}.
\end{align*}

## Simulation

### Set up

```{r}

# clear workspace
rm (list = ls(all=TRUE))

# set seed
set.seed(1234567, kind="Mersenne-Twister")

```

### Data Generating Process

\begin{align*}
y_t B + x_t A &= u_t \\
u_t &= u_{t-1} P + v_t \\
v_t &= N\left(0, V_t\right) \\
V_t &= S_t S_t^{'} \\
S_t &= C + D w_t \\
x_{1t} &\sim U\left[x_{1l},x_{1u}\right] \\
x_{1t} &\sim N \left(\mu_{x_{1}},\sigma_{x_{1}}^2\right)
\end{align*}

### Simulation

```{r}

# number of observations
t <- 2000

# parameters
b1 <- 0.6 
b2 <- 0.2 

a1 <- 0.4
a2 <- -0.5 

c11 <- 1.0
c21 <- 0.5
c22 <- 2.0

d11  <- 0.5
d21  <- 0.2
d22  <- 0.2

p11 <- 0.8 
p12 <- 0.1
p21 <- -0.2 
p22 <- 0.6

b  <-  matrix(c(1, -b2,
                -b1, 1), nrow=2, byrow=T)
a  <- matrix(c(-a1, 0,
               0, -a2), nrow=2, byrow=T)
c  <-  matrix(c(c11,  0,
                c21, c22), nrow=2, byrow=T)
d  <-  matrix(c(d11,  0,
                d21,  d22), nrow=2, byrow=T)

# exogenous variables                      
x <- cbind(10*runif(t), 3*rnorm(t))
w <- runif(t)

# disturbances 
zeros <- array(0, c(t,2))
u <- zeros
v <- zeros
for (i in 2:t) {
  l      <- c + d * w[i]  
  v[i,]  <- rnorm(2) %*% t(l)
  u[i,1] <- p11*u[i-1,1] + p12*u[i-1,2] + v[i,1]
  u[i,2] <- p21*u[i-1,1] + p22*u[i-1,2] + v[i,2]  
}

# simulate the reduced form   
y <- zeros
for (i in seq(t)) {
  y[i,] <- -x[i,] %*% a %*% solve(b) + u[i,] %*% solve(b)    
}

```

### Plots

```{r}

par(mfrow=c(1,2))

ts.plot(y[,1])
ts.plot(y[,2])

```

### Reduced form parameters

```{r}

# parameteres of the reduced form
ab <- -a %*% solve(b)
ab

```

## Least-Squares Estimator

Use the stacked model in matrix notation across equations $j$ and observations $i$.

```{r}

# dimensions
m <- ncol(y);m
t <- nrow(y);t
n <- t
k <- ncol(x);k

# stack regressands and regressors
Y <- as.vector(y) # stack y over observations
length(Y)

X <- diag(m) %x% x # stack x over observations
dim(X)

```

```{r}

# estimation
lm.res.01 <- lm(Y ~ X - 1)
lm.res.01$coefficients

# compare with reduced form parameters for simulation
as.vector(ab)

```

```{r}

# expand residuals again
e <- matrix(lm.res.01$residuals, ncol = m)
dim(e)

Sig.e <- 1/t * t(e) %*% e
Sig.e

# compare with reduced form parameters for simulation

```

Use the stacked model in matrix notation across observations $i$.

```{r}

# estimation
lm.res.02 <- lm(y ~ x - 1)
lm.res.02$coefficients

# compare with reduced form parameters for simulation
as.vector(ab)

```

## Sample Distribution of Least-Squares Estimator

<!-- .................................................. -->
<hr>

General distribution of $\widehat{\beta}$:

\begin{align*}
\sqrt{n} \left(\widehat{\beta} - \beta\right) &\underset{d}{\rightarrow}N\left(0, V_{\beta}\right)
\\
\widehat{\beta} &\underset{a}{\sim}N\left(\beta, \frac{V_{\beta}}{n}\right)
\end{align*}

General form:

\begin{align*}
V_{\beta} = Q^{-1} \Omega Q^{-1}
\end{align*}

<!-- .................................................. -->
<hr>

### Population Quantities

$\Omega$ in the population (robust w.r.t heteroskedasticity):

<div hidden>

\begin{align*}
\underset{\left(mk\times m\right)}{\overline{X}_{i}^{'}}\underset{\left(m\times 1\right)}{e_{i}} = \left(\begin{array}{c}
\underset{\left(k\times 1\right)}{x_{1i}}\underset{\left(k\times 1\right)}{e_{1i}} \\ \vdots \\ \underset{\left(k\times 1\right)}{x_{mi}}\underset{\left(1\times 1\right)}{e_{mi}} \end{array}\right)
\end{align*}

\begin{align*}
\frac{1}{\sqrt{n}}\sum_{i=1}^{n}\underset{\left(mk\times m\right)}{\overline{X}_{i}^{'}}\underset{\left(m\times 1\right)}{e_{i}} \underset{d}{\rightarrow}N\left(\underset{\left(mk\times 1\right)}{0}\underset{\left(mk\times mk\right)}{\Omega}\right)
\end{align*}

</div>

\begin{align*}
\underset{\left(mk\times mk\right)}{\Omega} &= \text{E}\left[\underset{\left(mk\times m\right)}{\overline{X}_{i}^{'}}\underset{\left(m\times 1\right)}{e_{i}^{\phantom{'}}}\underset{\left(1\times m\right)}{e_{i}^{'}}\underset{\left(m\times km\right)}{\overline{X}_{i}^{\phantom{'}}}\right]
\\
\phantom{\underset{\left(mk\times mk\right)}{\Omega}} &= \text{E} \left[ \underset{\left(mk \times m\right)}{\left(  \underset{\left(m \times m\right)}{I_{m}^{\phantom{'}}} \otimes \underset{\left(k \times 1\right)}{x_{i}^{\phantom{'}}} \right)} \underset{\left(m \times 1\right)}{e_{i}^{\phantom{'}}} \underset{\left(1 \times m\right)}{e_{i}^{'}} \underset{\left(m \times mk\right)}{\left( \underset{\left(m \times m\right)}{I_{m}^{\phantom{'}}} \otimes \underset{\left(1 \times k\right)}{x_{i}^{'}} \right)} \right]
\\
\phantom{\underset{\left(mk\times mk\right)}{\Omega}} &= \text{E}\left[\underset{\left(m \times 1\right)}{e_{i}^{\phantom{'}}}\underset{\left(1 \times m\right)}{e_{i}^{'}} \otimes \underset{\left(k \times 1\right)}{x_{i}^{\phantom{'}}}\underset{\left(1 \times k\right)}{x_{i}^{'}}\right]
\end{align*}

<!-- .................................................. -->
<hr>

$\Omega$ in the population (restricted to homoskedasticity):

\begin{align*}
\underset{\left(mk\times mk\right)}{\Omega} &= \text{E}\left[\underset{\left(mk\times m\right)}{\overline{X}_{i}^{'}}\underset{\left(m\times 1\right)}{e_{i}^{\phantom{'}}}\underset{\left(1\times m\right)}{e_{i}^{'}}\underset{\left(m\times km\right)}{\overline{X}_{i}^{\phantom{'}}}\right]
\\
\phantom{\underset{\left(mk\times mk\right)}{\Omega}} &= \text{E} \left[ \underset{\left(mk \times m\right)}{\left(  \underset{\left(m \times m\right)}{I_{m}^{\phantom{'}}} \otimes \underset{\left(k \times 1\right)}{x_{i}^{\phantom{'}}} \right)} \underset{\left(m \times m\right)}{\Sigma} \underset{\left(m \times mk\right)}{\left( \underset{\left(m \times m\right)}{I_{m}^{\phantom{'}}} \otimes \underset{\left(1 \times k\right)}{x_{i}^{'}} \right)} \right]
\\
\phantom{\underset{\left(mk\times mk\right)}{\Omega}} &= \text{E}\left[\underset{\left(m \times m\right)}{\Sigma} \otimes \underset{\left(k \times 1\right)}{x_{i}^{\phantom{'}}}\underset{\left(1 \times k\right)}{x_{i}^{'}}\right]
\end{align*}

with

\begin{align*}
\underset{\left(m \times m\right)}{\Sigma} = \text{E}\left[\underset{\left(m \times 1\right)}{e_{i}^{\phantom{'}}} \underset{\left(1 \times m\right)}{e_{i}^{'}} \right]
\end{align*}

<!-- .................................................. -->
<hr>

$\Omega$ in the population (robust w.r.t heteroskedasticity and serial correlation):

\begin{align*}
\underset{\left(mk\times mk\right)}{\Omega} &= \sum_{l=-\infty}^{\infty} \text{E}\left[\underset{\left(mk\times m\right)}{\overline{X}_{t-l}^{'}}\underset{\left(m\times 1\right)}{e_{t-l}^{\phantom{'}}}\underset{\left(1\times m\right)}{e_{t}^{'}}\underset{\left(m\times km\right)}{\overline{X}_{t}^{\phantom{'}}}\right]
\\
\phantom{\underset{\left(mk\times mk\right)}{\Omega}} &= \sum_{l=-\infty}^{\infty} \text{E} \left[ \underset{\left(mk \times m\right)}{\left(  \underset{\left(m \times m\right)}{I_{m}^{\phantom{'}}} \otimes \underset{\left(k \times 1\right)}{x_{t-l}^{\phantom{'}}} \right)} \underset{\left(m \times 1\right)}{e_{t-l}^{\phantom{'}}} \underset{\left(1 \times m\right)}{e_{t}^{'}} \underset{\left(m \times mk\right)}{\left( \underset{\left(m \times m\right)}{I_{m}^{\phantom{'}}} \otimes \underset{\left(1 \times k\right)}{x_{t}^{'}} \right)} \right]
\\
\phantom{\underset{\left(mk\times mk\right)}{\Omega}} &= \sum_{l=-\infty}^{\infty} \text{E}\left[\underset{\left(m \times 1\right)}{e_{t-l}^{\phantom{'}}}\underset{\left(1 \times m\right)}{e_{t}^{'}} \otimes \underset{\left(k \times 1\right)}{x_{t-l}^{\phantom{'}}}\underset{\left(1 \times k\right)}{x_{t}^{'}}\right]
\end{align*}

<!-- .................................................. -->
<hr>

$Q$ in the population:

\begin{align*}
\underset{\left(mk\times mk\right)}{Q} &= \text{E} \left[ \overline{X}_{i}^{'} \overline{X}_{i} \right]
\\
\phantom{Q} &= \text{E} \left[ \left( I_{m} \otimes x_{i} \right) \left( I_{m} \otimes x_{i}^{'} \right) \right]
\\
\phantom{Q} &= I_{m} \otimes \text{E} \left[ x_{i}^{'} x_{i} \right]
\end{align*}

<!-- .................................................. -->
<hr>

### Sample Quantities

Sample counterpart of $\Omega$ (robust w.r.t heteroskedasticity):

<!-- \begin{align*} -->
<!-- \underset{\left(mk\times mk\right)}{\widehat{\Omega}} &= \sum_{i=1}^{n} \underset{\left(mk\times m\right)}{\overline{X}_{i}^{'}}\underset{\left(m\times 1\right)}{\widehat{e}_{i}}\underset{\left(1\times m\right)}{\widehat{e}_{i}^{'}}\underset{\left(m\times km\right)}{\overline{X}_{i}} -->
<!-- \\ -->
<!-- \phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \sum_{i=1}^{n} \left( \left( I_{m} \otimes x_{i} \right) \widehat{e}_{i} \widehat{e}_{i}^{'} \left( I_{m} \otimes x_{i}^{'} \right) \right) -->
<!-- \\ -->
<!-- \phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \sum_{i=1}^{n}\left(\widehat{e}_{i}\widehat{e}_{i}^{'} \otimes x_{i}x_{i}^{'}\right) -->
<!-- \end{align*} -->

\begin{align*}
\underset{\left(mk\times mk\right)}{\widehat{\Omega}} &= \frac{1}{n^2} \sum_{i=1}^{n} \underset{\left(mk\times m\right)}{\overline{X}_{i}^{'}}\underset{\left(m\times 1\right)}{\widehat{e}_{i}}\underset{\left(1\times m\right)}{\widehat{e}_{i}^{'}}\underset{\left(m\times km\right)}{\overline{X}_{i}}
\\
\phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \frac{1}{n^2} \sum_{i=1}^{n} \left( \left( I_{m} \otimes x_{i} \right) \widehat{e}_{i} \widehat{e}_{i}^{'} \left( I_{m} \otimes x_{i}^{'} \right) \right)
\\
\phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \frac{1}{n^2} \sum_{i=1}^{n}\left(\widehat{e}_{i}\widehat{e}_{i}^{'} \otimes x_{i}x_{i}^{'}\right)
\end{align*}

```{r, include=FALSE}

Omega.02 <- matrix(0, nrow = m*k, ncol = m*k)
for (i in 1:t) {
  Omega.02 <- Omega.02 + ((t(e[i,,drop=F]) %*% e[i,,drop=F]) %x% (t(x[i,,drop=F]) %*% x[i,,drop=F]))
}
Omega.02

1/n * Omega.02

```

```{r}

xe.01 <- matrix(seq(1,t*m*k), nrow = t, ncol = m*k)
for (i in 1:t) {
  for (j in 1:m) {
    xe.01[i,(k*(j-1)+1):(k*j)] <- x[i,,drop=F] * e[i,j,drop=T]
  }
}

Omega.02.hat <- 1/(n^2) * (t(xe.01) %*% xe.01)
Omega.02.hat

```

Compare with the results of the `sandwich` package.

```{r}

Omega.02.hat * n

sandwich::meat(lm.res.02)

```

```{r, include=FALSE}

head(xe.01)
tail(xe.01)

head(sandwich::estfun(lm.res.02))
tail(sandwich::estfun(lm.res.02))
# seems to work

Omega.02.hat

sandwich::meat(lm.res.02)

```

<!-- .................................................. -->
<hr>

Sample counterpart of $\Omega$ (restricted to homoskedasticity):

<!-- \begin{align*} -->
<!-- \underset{\left(mk\times mk\right)}{\widehat{\Omega}} &= \sum_{i=1}^{n} \underset{\left(mk\times m\right)}{\overline{X}_{i}^{'}}\underset{\left(m\times 1\right)}{\widehat{e}_{i}}\underset{\left(1\times m\right)}{\widehat{e}_{i}^{'}}\underset{\left(m\times km\right)}{\overline{X}_{i}} -->
<!-- \\ -->
<!-- \phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \sum_{i=1}^{n} \left( \left( I_{m} \otimes x_{i} \right) \widehat{\Sigma} \left( I_{m} \otimes x_{i}^{'} \right) \right) -->
<!-- \\ -->
<!-- \phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \sum_{i=1}^{n} \left( \widehat{\Sigma} \otimes x_{i}^{\phantom{'}}x_{i}^{'} \right) -->
<!-- \\ -->
<!-- \phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \widehat{\Sigma} \otimes \sum_{i=1}^{n}  x_{i}^{\phantom{'}}x_{i}^{'} -->
<!-- \end{align*} -->

<!-- with -->

<!-- \begin{align*} -->
<!-- \widehat{\Sigma} = \frac{1}{n} \sum_{i=1}^{n} \left( {e_{i} e_{i}^{'}} \right) -->
<!-- \end{align*} -->

\begin{align*}
\underset{\left(mk\times mk\right)}{\widehat{\Omega}} &= \frac{1}{n^2} \sum_{i=1}^{n} \underset{\left(mk\times m\right)}{\overline{X}_{i}^{'}}\underset{\left(m\times 1\right)}{\widehat{e}_{i}}\underset{\left(1\times m\right)}{\widehat{e}_{i}^{'}}\underset{\left(m\times km\right)}{\overline{X}_{i}}
\\
\phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \frac{1}{n^2} \sum_{i=1}^{n} \left( \left( I_{m} \otimes x_{i} \right) \widehat{\Sigma} \left( I_{m} \otimes x_{i}^{'} \right) \right)
\\
\phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \frac{1}{n^2} \sum_{i=1}^{n} \left( \widehat{\Sigma} \otimes x_{i}^{\phantom{'}}x_{i}^{'} \right)
\\
\phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \frac{1}{n^2} \left( \widehat{\Sigma} \otimes \sum_{i=1}^{n}  x_{i}^{\phantom{'}}x_{i}^{'} \right)
\end{align*}

with

\begin{align*}
\widehat{\Sigma} = \frac{1}{n} \sum_{i=1}^{n} \left( {e_{i} e_{i}^{'}} \right)
\end{align*}

```{r}

Omega.01.hat <- 1/(n^2) * ((t(e) %*% e) / n) %x% (t(x) %*% x)
Omega.01.hat

```

<!-- .................................................. -->
<hr>

Sample counterpart of $\Omega$ (robust w.r.t heteroskedasticity and serial correlation):

```{r}

library(sandwich)

# sandwich::vcovHAC(lm.res.02)
# # Error in umat - res : non-conformable arrays

sandwich___bwAndrews <- function (x, order.by = NULL,
                                  kernel = c("Quadratic Spectral", "Truncated", "Bartlett", "Parzen", "Tukey-Hanning"),
                                  approx = c("AR(1)", "ARMA(1,1)"), weights = NULL, prewhite = 1, ar.method = "ols", data = list(), ...) {
  
  # inputs
  x <- lm.res.02
  order.by <- NULL
  kernel <- c("Quadratic Spectral")
  approx <- c("AR(1)")
  weights <- NULL
  prewhite <- 1
  ar.method <- "ols"
  data = list()

  
    
  if (is.list(x) && !is.null(x$na.action)) {
    class(x$na.action) <- "omit"
  }
  kernel <- match.arg(kernel)
  approx <- match.arg(approx)
  prewhite <- as.integer(prewhite)
  umat <- if (inherits(x, "matrix")) {
    x
  } else {
    estfun(x)[, , drop = FALSE]
  }
  if (zoo::is.zoo(umat)) {
    umat <- as.matrix(coredata(umat))
  }
  n <- nrow(umat)
  k <- ncol(umat)
  if (!is.null(order.by)) {
    if (inherits(order.by, "formula")) {
      z <- model.matrix(order.by, data = data)
      z <- as.vector(z[, ncol(z)])
    } else {
      z <- order.by
    }
    index <- order(z)
  } else {
    index <- 1:n
  }
  umat <- umat[index, , drop = FALSE]
  
  # problem umat (n x m*k) - res ()
  if (is.null(weights)) {
    weights <- rep(1, k)
    unames <- colnames(umat)
    if (!is.null(unames) && "(Intercept)" %in% unames) {
      weights[which(unames == "(Intercept)")] <- 0
    } else {
      # res <- try(as.vector(rowMeans(estfun(x)/model.matrix(x), na.rm = TRUE)), silent = TRUE)
      # MM: added for m=2!
      tmp <- model.matrix(x)
      for (i in 1:ncol(residuals(x))) { tmp <- cbind(tmp, model.matrix(x)) }
      res <- try(as.vector(rowMeans(estfun(x)/cbind(model.matrix(x), model.matrix(x)), na.rm = TRUE)), silent = TRUE)
      if (inherits(res, "try-error")) {
        res <- try(residuals(x), silent = TRUE)
      }
      if (!inherits(res, "try-error")) {
        weights[which(colSums((umat - res)^2) < 1e-16)] <- 0
      }
    }
    if (isTRUE(all.equal(weights, rep(0, k)))) {
      weights <- rep(1, k)
    }
  } else {
    weights <- rep(weights, length.out = k)
  }
  if (length(weights) < 2) {
    weights <- 1
  }
  if (prewhite > 0) {
    var.fit <- ar(umat, order.max = prewhite, demean = FALSE, aic = FALSE, method = ar.method)
    if (inherits(var.fit, "try-error")) {
      stop(sprintf("VAR(%i) prewhitening of estimating functions failed", prewhite))
    }
    umat <- as.matrix(na.omit(var.fit$resid))
    n <- n - prewhite
  }
  if (approx == "AR(1)") {
    fitAR1 <- function(x) {
      rval <- ar(x, order.max = 1, aic = FALSE, method = "ols")
      rval <- c(rval$ar, sqrt(rval$var.pred))
      names(rval) <- c("rho", "sigma")
      return(rval)
    }
    ar.coef <- apply(umat, 2, fitAR1)
    denum <- sum(weights * (ar.coef["sigma", ]/(1 - ar.coef["rho", ]))^4)
    alpha2 <- sum(weights * 4 * ar.coef["rho", ]^2 * ar.coef["sigma", ]^4/(1 - ar.coef["rho", ])^8)/denum
    alpha1 <- sum(weights * 4 * ar.coef["rho", ]^2 * ar.coef["sigma", ]^4/((1 - ar.coef["rho", ])^6 * (1 + ar.coef["rho", ])^2))/denum
  } else {
    fitARMA11 <- function(x) {
      rval <- arima(x, order = c(1, 0, 1), include.mean = FALSE)
      rval <- c(rval$coef, sqrt(rval$sigma2))
      names(rval) <- c("rho", "psi", "sigma")
      return(rval)
    }
    arma.coef <- apply(umat, 2, fitARMA11)
    denum <- sum(weights * ((1 + arma.coef["psi", ]) * arma.coef["sigma", ]/(1 - arma.coef["rho", ]))^4)
    alpha2 <- sum(weights * 4 * ((1 + arma.coef["rho", ] * arma.coef["psi", ]) * (arma.coef["rho", ] + arma.coef["psi", ]))^2 * arma.coef["sigma", ]^4/(1 - arma.coef["rho", ])^8)/denum
    alpha1 <- sum(weights * 4 * ((1 + arma.coef["rho", ] * arma.coef["psi", ]) * (arma.coef["rho", ] + arma.coef["psi", ]))^2 * arma.coef["sigma", ]^4/((1 - arma.coef["rho", ])^6 * (1 + arma.coef["rho", ])^2))/denum
  }
  
  rval <- switch(kernel,
                 Truncated = { 0.6611 * (n * alpha2)^(1/5) },
                 Bartlett = { 1.1447 * (n * alpha1)^(1/3) },
                 Parzen = { 2.6614 * (n * alpha2)^(1/5) },
                 `Tukey-Hanning` = { 1.7462 * (n * alpha2)^(1/5) },
                 `Quadratic Spectral` = { 1.3221 * (n * alpha2)^(1/5) } )
  
  return(rval)
    
}

sandwich__kweights <- function (x,
                                kernel = c("Truncated", "Bartlett", "Parzen", "Tukey-Hanning", "Quadratic Spectral"),
                                normalize = FALSE) {
  
  kernel <- match.arg(kernel, c("Truncated", "Bartlett", "Parzen", "Tukey-Hanning", "Quadratic Spectral"))
  if (normalize) {
    ca <- switch(kernel,
                 Truncated = 2,
                 Bartlett = 2/3,
                 Parzen = 0.539285, 
                 `Tukey-Hanning` = 3/4,
                 `Quadratic Spectral` = 1)
  } else {
    ca <- 1
  }
  switch(kernel,
         Truncated = { ifelse(ca * abs(x) > 1, 0, 1) },
         Bartlett = { ifelse(ca * abs(x) > 1, 0, 1 - abs(ca * x)) },
         Parzen = { ifelse(ca * abs(x) > 1, 0, ifelse(ca * abs(x) < 0.5, 1 - 6 * (ca * x)^2 + 6 * abs(ca * x)^3, 2 * (1 - abs(ca * x))^3)) },
         `Tukey-Hanning` = { ifelse(ca * abs(x) > 1, 0, (1 + cos(pi * ca * x))/2) },
         `Quadratic Spectral` = {
           qs <- function(x) {
             y <- 6 * pi * x/5
             3 * (1/y)^2 * (sin(y)/y - cos(y))
            }
          w <- qs(x)
          if (length(ix <- which(abs(x) < 0.001)) > 0L) {
            cf <- 1e+06 * log(qs(0.001))
            w[ix] <- exp(cf * x[ix]^2)
          }
          w}
         )
}

sandwich__weightsAndrews <- function (x, order.by = NULL, bw = bwAndrews,
                                      kernel = c("Quadratic Spectral", "Truncated", "Bartlett", "Parzen", "Tukey-Hanning"),
                                      prewhite = 1, ar.method = "ols", tol = 1e-07, data = list(), verbose = FALSE, ...) {
  
  # # inputs
  # x <- lm.res.02
  # order.by <- NULL
  # bw <- bwAndrews
  # kernel <- NULL
  # prewhite <- 1
  # ar.method <- "ols"
  # tol <- 1e-07
  # verbose <- FALSE
  # data = list()

  
  
  if (is.list(x) && !is.null(x$na.action)) {
    class(x$na.action) <- "omit"
  }
  kernel <- match.arg(kernel)
  
  # call bandwidth function bw(), bwAndrews()
  if (is.function(bw)) {
    # bw <- bw(x, order.by = order.by, kernel = kernel, prewhite = prewhite, data = data, ar.method = ar.method, ...)
    # bw <- bw(x, order.by = order.by, kernel = kernel, prewhite = prewhite, data = data, ar.method = ar.method)
    bw <- sandwich___bwAndrews(x, order.by = order.by, kernel = kernel, prewhite = prewhite, data = data, ar.method = ar.method)
  }
      
  if (verbose) {
    cat(paste("\nBandwidth chosen:", format(bw), "\n"))
  }
  
  n <- NROW(estfun(x)) - as.integer(prewhite)
  # call kernel weights function kweights(), weightsAndrews()
  # weights <- kweights(0:(n - 1)/bw, kernel = kernel)
  weights <- sandwich__kweights(0:(n - 1)/bw, kernel = kernel)
  weights <- weights[1:max(which(abs(weights) > tol))]
  
  return(weights)
    
}

sandwich__meatHAC <- function(x, order.by = NULL, prewhite = FALSE, weights = weightsAndrews, 
                               adjust = TRUE, diagnostics = FALSE, ar.method = "ols", 
                               data = list(), ...) {
  
  # # inputs
  # x <- lm.res.02
  # order.by <- NULL
  # prewhite <- FALSE
  # weights <- weightsAndrews 
  # adjust <- TRUE
  # diagnostics <- FALSE
  # ar.method <- "ols"
  # data <- list()

  
    
  if (is.list(x) && !is.null(x$na.action)) {
    class(x$na.action) <- "omit"
  }
  prewhite <- as.integer(prewhite)
  # umat <- estfun(x, ...)[, , drop = FALSE]
  umat <- estfun(x)[, , drop = FALSE]
  if (zoo::is.zoo(umat)) {
    umat <- as.matrix(coredata(umat))
  }
  # MM: check
  head(xe.01)
  head(umat) # (t x m*k)
  
  n.orig <- n <- nrow(umat)
  k <- ncol(umat)
  if (!is.null(order.by)) {
    if (inherits(order.by, "formula")) {
      z <- model.matrix(order.by, data = data)
      z <- as.vector(z[, ncol(z)])
    } else {
      z <- order.by
    }
    index <- order(z)
  } else {
    index <- 1:n
  }
  umat <- umat[index, , drop = FALSE]
  if (prewhite > 0) {
    var.fit <- try(ar(umat, order.max = prewhite, demean = FALSE, aic = FALSE, method = ar.method))
    if (inherits(var.fit, "try-error")) {
      stop(sprintf("VAR(%i) prewhitening of estimating functions failed", prewhite))
    }
    if (k > 1) {
      D <- solve(diag(ncol(umat)) - apply(var.fit$ar, 2:3, sum))
    } else {
      D <- as.matrix(1/(1 - sum(var.fit$ar)))
    }
    umat <- as.matrix(na.omit(var.fit$resid))
    n <- n - prewhite
  }
  
  # call weight function weights(), weightsAndrews()
  if (is.function(weights)) {
    # weights <- weights(x, order.by = order.by, prewhite = prewhite, ar.method = ar.method, data = data)
    weights <- sandwich__weightsAndrews(x, order.by = order.by, prewhite = prewhite, ar.method = ar.method, data = data)
  }
  if (length(weights) > n) {
    warning("more weights than observations, only first n used")
    weights <- weights[1:n]
  }
  utu <- 0.5 * crossprod(umat) * weights[1]
  wsum <- n * weights[1]/2
  w2sum <- n * weights[1]^2/2
  if (length(weights) > 1) {
    for (ii in 2:length(weights)) {
      utu <- utu + weights[ii] * crossprod(umat[1:(n - ii + 1), , drop = FALSE], umat[ii:n, , drop = FALSE])
      wsum <- wsum + (n - ii + 1) * weights[ii]
      w2sum <- w2sum + (n - ii + 1) * weights[ii]^2
    }
  }
  utu <- utu + t(utu)
  if (adjust) {
    utu <- n.orig/(n.orig - k) * utu
  }
  if (prewhite > 0) {
    utu <- crossprod(t(D), utu) %*% t(D)
  }
  wsum <- 2 * wsum
  w2sum <- 2 * w2sum
  bc <- n^2/(n^2 - wsum)
  df <- n^2/w2sum
  rval <- utu/n.orig
  if (diagnostics) {
    attr(rval, "diagnostics") <- list(bias.correction = bc, df = df)
  }
    
  return(rval)
  
}

tmp <- sandwich__meatHAC(x = lm.res.02, order.by = NULL, prewhite = FALSE,
                  weights = weightsAndrews,
                  adjust = TRUE, diagnostics = FALSE, ar.method = "ols", data = list())

sandwich::sandwich(lm.res.02, meat. = tmp)

vcovHC(lm.res.02)

```





<!-- .................................................. -->
<hr>

Sample counterpart of $Q$:

<!-- \begin{align*} -->
<!-- \underset{\left(mk\times mk\right)}{Q} &= \sum_{i=1}^{n} \left( \overline{X}_{i}^{'} \overline{X}_{i} \right) -->
<!-- \\ -->
<!-- \phantom{Q} &= \sum_{i=1}^{n} \left( \left( I_{m} \otimes x_{i} \right) \left( I_{m} \otimes x_{i}^{'} \right) \right) -->
<!-- \\ -->
<!-- \phantom{Q} &= I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right) -->
<!-- \end{align*} -->

\begin{align*}
\underset{\left(mk\times mk\right)}{Q} &= \frac{1}{n} \sum_{i=1}^{n} \left( \overline{X}_{i}^{'} \overline{X}_{i} \right)
\\
\phantom{Q} &= \frac{1}{n} \sum_{i=1}^{n} \left( \left( I_{m} \otimes x_{i} \right) \left( I_{m} \otimes x_{i}^{'} \right) \right)
\\
\phantom{Q} &= \frac{1}{n} \left( I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right) \right)
\end{align*}

```{r}

Q.hat <- 1/n * diag(m) %x% (t(x) %*% x)
Q.hat

```

Compare with the results of the `sandwich` package.

```{r}

solve(Q.hat)

sandwich::bread(lm.res.02)

```



<!-- .................................................. -->
<hr>

Sample counterpart of $V_{\beta}$ (robust w.r.t heteroskedasticity):

<!-- \begin{align*} -->
<!-- \widehat{V}_{\widehat{\beta}} &= \widehat{Q}^{-1} \widehat{\Omega} \widehat{Q}^{-1} -->
<!-- \\ -->
<!-- \phantom{\widehat{V}_{\widehat{\beta}}} &= \left(I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right)\right)^{-1} \sum_{i=1}^{n}\left(\widehat{e}_{i}\widehat{e}_{i}^{'} \otimes x_{i}x_{i}^{'}\right) \left(I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right)\right)^{-1} -->
<!-- \\ -->
<!-- \phantom{\widehat{V}_{\widehat{\beta}}} &= \left(I_{m} \otimes \left( X^{'} X \right)^{-1}\right) \sum_{i=1}^{n}\left(\widehat{e}_{i}\widehat{e}_{i}^{'} \otimes x_{i}x_{i}^{'}\right) \left(I_{m} \otimes \left( X^{'} X \right)^{-1}\right) -->
<!-- \end{align*} -->

\begin{align*}
\widehat{V}_{\widehat{\beta}} &= \widehat{Q}^{-1} \widehat{\Omega} \widehat{Q}^{-1}
\\
\phantom{\widehat{V}_{\widehat{\beta}}} &= \left(\frac{1}{n} \left( I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right) \right)\right)^{-1} \frac{1}{n^2} \sum_{i=1}^{n}\left(\widehat{e}_{i}\widehat{e}_{i}^{'} \otimes x_{i}x_{i}^{'}\right) \left(\frac{1}{n} \left( I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right) \right)\right)^{-1}
\end{align*}

<!-- \phantom{\widehat{V}_{\widehat{\beta}}} &= \left(I_{m} \otimes \left( \frac{1}{n} X^{'} X \right)^{-1}\right) \frac{1}{n^2} \sum_{i=1}^{n}\left(\widehat{e}_{i}\widehat{e}_{i}^{'} \otimes x_{i}x_{i}^{'}\right) \left(I_{m} \otimes \left( \frac{1}{n} X^{'} X \right)^{-1}\right) -->

```{r}

V.b.hat.02 <- solve(Q.hat) %*% Omega.02.hat %*% solve(Q.hat)
V.b.hat.02

```

Compare with the results of the `sandwich` package.

```{r}

V.b.hat.02

sandwich::vcovHC(lm.res.02, type = "HC0")

```

<!-- .................................................. -->
<hr>

Sample counterpart of $V_{\beta}$ (restricted to homoskedasticity):

<!-- \begin{align*} -->
<!-- \widehat{V}_{\widehat{\beta}} &= \widehat{Q}^{-1} \widehat{\Omega} \widehat{Q}^{-1} -->
<!-- \\ -->
<!-- \phantom{\widehat{V}_{\widehat{\beta}}} &= \left(I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right)\right)^{-1} \sum_{i=1}^{n} \left( \left( I_{m} \otimes x_{i} \right) \widehat{\Sigma} \left( I_{m} \otimes x_{i}^{'} \right) \right) \left(I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right)\right)^{-1} -->
<!-- \\ -->
<!-- \phantom{\widehat{V}_{\widehat{\beta}}}  -->
<!-- \phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \widehat{\Sigma} \otimes \sum_{i=1}^{n}  x_{i}^{\phantom{'}}x_{i}^{'} -->
<!-- \end{align*} -->

\begin{align*}
\widehat{V}_{\widehat{\beta}} &= \widehat{Q}^{-1} \widehat{\Omega} \widehat{Q}^{-1}
\\
\phantom{\widehat{V}_{\widehat{\beta}}} &= \left(\frac{1}{n} \left( I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right) \right)\right)^{-1} \frac{1}{n^2} \sum_{i=1}^{n} \left( \left( I_{m} \otimes x_{i} \right) \widehat{\Sigma} \left( I_{m} \otimes x_{i}^{'} \right) \right) \left(\frac{1}{n} \left( I_{m} \otimes \sum_{i=1}^{n} \left( x_{i}^{'} x_{i} \right) \right)\right)^{-1}
\\
\phantom{\widehat{V}_{\widehat{\beta}}} 
\phantom{\underset{\left(mk\times mk\right)}{\widehat{\Omega}}} &= \frac{1}{n^2} \left( \widehat{\Sigma} \otimes \sum_{i=1}^{n}  x_{i}^{\phantom{'}}x_{i}^{'} \right) 
\end{align*}

```{r}

V.b.hat.01 <- solve(Q.hat) %*% Omega.01.hat %*% solve(Q.hat)
V.b.hat.01

```

Compare with results of the `stats` package.

```{r}

V.b.hat.01 * n/(n-2)

vcov(lm.res.02)

class(lm.res.02)

```

<!-- .................................................. -->
<hr>

Sample counterpart of $V_{\beta}$ (robust w.r.t heteroskedasticity and serial correlation):

...





```{r, include=FALSE, eval=TRUE}

# some checks

# approach by Hansen and Seo (2002)

xx <- solve(t(x) %*% x)
xxx <- x %*% xx
u <- y - x %*% xx %*% (t(x) %*% y)

head(u) # seems to work!
head(e)

# # 2nd-stage cointegration (restricted)
# w0 <- as.matrix(xlag%*%rbind(1,-b0))
# z0 <- cbind(w0,x)

# 2nd-stage unrestricted
z0 <- x

kk <- ncol(z0)
zz0 <- solve(t(z0)%*%z0)
zzz0 <- z0%*%zz0
beta0 <- t(zzz0)%*%y
e <- y - z0%*%beta0 # note, for restricted same as u above
sige <- t(e)%*%e/t
nlike <- (t/2)*log(det(sige))

cov <- 1

k_product <- function(ma,mb) {
  mat <- matrix(0, nrow(ma)*nrow(mb), ncol(ma)*ncol(mb))
  for (i in 1:nrow(ma)) {
    for (j in 1:ncol(ma)) {
      mat[(1:nrow(mb))+(i-1)*nrow(mb),(1:ncol(mb))+(j-1)*ncol(mb)] <- mb*ma[i,j]
    }
  }
  mat
}

if (cov == 1){
  xe <- cbind((z0*(e[,1] %*% matrix(1,1,ncol(z0)))), (z0*(e[,2] %*% matrix(1,1,ncol(z0)))))
  m0 <- k_product(diag(2), zz0)
  v <- m0 %*% t(xe) %*% xe %*% m0
}else{
  v <- k_product(sige, zz0)
}
se <- as.matrix(sqrt(diag(v)))

v

V.b.hat.02
# seems to work!

t(xe) %*% xe
Omega.02
# seems to work!

m0
solve(Q.hat)
solve(Q.hat * n)
# seems to work!

# non robust
k_product(sige, zz0)
V.b.hat.01
(t(e) %*% e/t) %x% solve(t(x) %*% x)
# seems to work!

library(sandwich)

lm.res.03 <- lm(y ~ x - 1)

bread(lm.res.03) %*% meat(lm.res.03) %*% bread(lm.res.03) / n
sandwich::vcovHC(lm.res.03, type = "HC0")
v
V.b.hat.02
# seems to work!

```



```{r, include=FALSE, eval=FALSE}

library(MASS)

lm.res <- lm(Gas ~ Temp + 1, data = whiteside)

lm.sum.res <- summary(lm.res)

# START

y <- as.matrix(whiteside$Gas)
n <- nrow(y)
m <- ncol(y)
x <- cbind(as.matrix(whiteside$Temp), as.matrix(rep(1,n)))
k <- ncol(x)

# stack regressands and regressors
Y <- as.vector(y) # stack y over observations
length(Y)

X <- diag(m) %x% x # stack x over observations
dim(X)



# estimation
lm.res.x <- lm(Y ~ X - 1)
lm.res.x$coefficients

# expand residuals again
e <- matrix(lm.res$residuals, ncol = m)
dim(e)

Sig.e <- (t(e) %*% e)/n
Sig.e

Omega.01 <- ((t(e) %*% e)/n) %x% (t(x) %*% x)
Omega.01

Omega.01.hat <- 1/n * Omega.01
Omega.01.hat

Q <- diag(m) %x% (t(x) %*% x)
Q

Q.hat <- 1/n * Q
Q.hat

V.b.hat.01 <- solve(Q.hat) %*% Omega.01.hat %*% solve(Q.hat)
V.b.hat.01

vcov(lm.res)

solve(lm.sum.res$cov.unscaled)
Q.hat

crossprod(x)

```

