---
output: html_document
---

# Topic 04

```{r,include=FALSE}

library(knitr)
knitr::opts_chunk$set(
  include = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = TRUE
  # fig.height = 8,
  # fig.width = 10,
  # fig.align = 'center'
  )
# options(scipen=1)
# options(width=9999)
# # See: https://stackoverflow.com/questions/36230790/missing-horizontal-scroll-bar-in-r-markdown-html-code-chunks-and-output

```

Complete model:

\begin{align*}
\Delta y_{t} =\mu_{0} + \mu_{1}t + \alpha\beta^{'}y_{t-1}
\end{align*}

Deterministic terms:

\begin{align*}
\mu_{j} &= \delta_{j} + \alpha\beta_{j}^{'}
\\
\underset{\left(N\times 1\right)}{\mu_{j}} &= \underset{\left(N\times 1\right)}{\delta_{j}} + \underset{\left(N\times r\right)}{\alpha}\underset{\left(r\times 1\right)}{\beta^{'}_{j}}
\end{align*}



Simulation

Model 1

* No deterministic terms

...

Model 2

* Constant restricted to the cointegrating space only
* [See with some adjustments](https://www.pfaffikus.de/books/spex2/files/Rcode-4-3.R)

```{r}

library(urca)
set.seed(12345)
e1 <- rnorm(250, 0, 0.5)
e2 <- rnorm(250, 0, 0.5)
e3 <- rnorm(250, 0, 0.5)
u1.ar1 <- arima.sim(model = list(ar = 0.75),
                    innov = e1, n = 250)
u2.ar1 <- arima.sim(model = list(ar = 0.3),
                    innov = e2, n = 250)
y3 <- cumsum(e3)
y1 <- 0.1 + 0.8 * y3 + u1.ar1
y2 <- 0.2 - 0.3 * y3 + u2.ar1
y.mat <- data.frame(y1, y2, y3)
vecm <- ca.jo(y.mat)
jo.results <- summary(vecm)
vecm.r2 <- cajorls(vecm, r = 2)
class(jo.results)
slotNames(jo.results)

```

```{r}

plot(seq(1, nrow(y.mat)),
     y1,
     col = "blue",
     type = "l",
     ylim = c(min(y1, y2, y3), max(y1, y2, y3)))
lines(seq(1, nrow(y.mat)),
      y2,
      col = "red",
      type = "l")
lines(seq(1, nrow(y.mat)),
      y3,
      col = "darkgreen",
      type = "l")


```

Model 3

* Unrestricted constant
* MHH, 2011, (p. 767)

```{r}

library(MASS)

# Specify model
beta <- c(1,-1)                 # Normalized cointegrating vector                             
beta0 <- 2                      # Long-run constant                                          
beta1 <- 0                      # Long-run time trend                                        

alpha <- c(-0.1,0.1)            # Error-correction parameters                               

alpha0 <- Null(alpha %*% t(alpha))
delta0 <- alpha0*2              # Short-run intercepts using orthogonal complement of alpha   
delta1 <- alpha0*0              # Short-run time trend using orthogonal complement of alpha 

# Simulate model
t <- 200
y <- array(0, c(t,2))
v <- array(rnorm(t*2), c(t,2))

for (j in 2:t) {
  u <- beta0 + beta1*j + y[j-1,] %*% beta
  y[j,] <- y[j-1,] + t(delta0 + delta1*j + alpha*u) + v[j,]
}

```

```{r}

plot(seq(1, nrow(y)),
     y[,1],
     col = "blue",
     type = "l",
     ylim = c(min(y), max(y)))
lines(seq(1, nrow(y)),
      y[,2],
      col = "red",
      type = "l")

```

\begin{align*}
\delta_{j} = \alpha_{\perp}\left(\alpha_{\perp}^{'}\alpha_{\perp}\right)^{-1}\alpha_{\perp}^{'}\mu_{j}
\end{align'}


