---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Topic 05

```{r,include=FALSE}

library(knitr)
knitr::opts_chunk$set(
  include = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = TRUE,
  cache.path = "./cache-files"
  # fig.height = 8,
  # fig.width = 10,
  # fig.align = 'center'
  )
# options(scipen=1)
# options(width=9999)
# # See: https://stackoverflow.com/questions/36230790/missing-horizontal-scroll-bar-in-r-markdown-html-code-chunks-and-output

```

## Motivation

Comparison of the analysis of multivariate time series using the `urca` in combination with the `vars` package and the `tsDyn` package

## Linear VECM model estimation

```{r}

plot(tsDyn::zeroyldMeta$Date, tsDyn::zeroyldMeta$short.run, type = "l",
     xlab = "Date", ylab = "Rate")
lines(tsDyn::zeroyldMeta$Date, tsDyn::zeroyldMeta$long.run, lty = 2)
legend("topleft", lty = c(1, 2),
       legend = c("Short rate: 12 months", "Long rate: 120 months"))

```

```{r}

vecm.tsdyn <- tsDyn::VECM(tsDyn::zeroyldMeta[, c("long.run", "short.run")], r = 1, lag = 1, estim = "ML")
summary(vecm.tsdyn)

```

```{r}

vecm.urca <- urca::ca.jo(x = tsDyn::zeroyldMeta[, c("long.run", "short.run")], ecdet = "none", K = 2)
vecm.urca@V

```

## Linear VECM model orthogonalized IRF analysis

Compare non-orthogonalized IRFs first

```{r}

irf.tsdyn <- tsDyn:::irf.VECM(vecm.tsdyn, ortho = FALSE, boot = FALSE)

irf.tsdyn$irf$short.run
irf.tsdyn$irf$long.run

```

```{r}

v2v.vars <- vars::vec2var(vecm.urca, r = 1)
v2v.vars <- vars::irf(v2v.vars, ortho = FALSE, boot = FALSE)

v2v.vars$irf$short.run
v2v.vars$irf$long.run

```

Compare orthongonalized IRFs afterwards

```{r}

irf.tsdyn <- tsDyn:::irf.VECM(vecm.tsdyn, ortho = TRUE, boot = FALSE)

irf.tsdyn$irf$short.run
irf.tsdyn$irf$long.run

```

```{r}

v2v.vars <- vars::vec2var(vecm.urca, r = 1)
v2v.vars <- vars::irf(v2v.vars, ortho = TRUE, boot = FALSE)

v2v.vars$irf$short.run
v2v.vars$irf$long.run

```

## Linear VECM model generalized IRF analysis

Step x: Check class and `GIRF` method of linear object

```{r}

class(vecm.tsdyn)

tsDyn:::GIRF.nlVar

```

Step x: Call function `tsDyn:::GIRF.setar()`

```{r, cache = TRUE}

# tsDyn_GIRF_setar_fun no 01 ----

# tsDyn_GIRF_setar_fun - function (object, n.ahead = 10, seed = NULL, n.hist = 20, n.shock = 20, 
#                                  R = 10, hist_li = NULL, shock_li = NULL, ...) {
  
  # inputs
  object <- vecm.tsdyn
  n.ahead <- 10
  seed <- NULL
  n.hist = 20
  n.shock = 20
  R = 10
  hist_li <- NULL
  shock_li <- NULL
  
    
  
  # lag <- tsDyn:::get_lag(object)
  lag <- tsDyn:::get_lag.nlVar(object)
  n_start <- ifelse(inherits(object, c("VECM", "TVECM")), lag + 1, lag)
  # x_orig <- get_data_orig(object, as.df = TRUE)
  x_orig <- tsDyn:::get_data_orig.nlVar(object, as.df = TRUE)
  N <- nrow(x_orig)
  resids <- as.data.frame(residuals(object, initVal = FALSE))
  n_used <- nrow(resids)
  # nVar <- get_nVar(object)
  nVar <- tsDyn:::get_nVar.nlVar(object)
  
  sample_hist <- function() {
    hist_M <- sample(n_start:N, size = 1, replace = FALSE)
    (hist_M - n_start + 1):hist_M
  }
  
  nrow_length <- function(x) {
    nr <- nrow(x)
    if (is.null(nr)) {
      nr <- length(x)
    }
    nr
  }
  
  if (is.null(hist_li)) {
    if (!is.null(seed)) {
      set.seed(seed)
    }
    samples_hist <- sample(n_start:N, size = n.hist, replace = FALSE)
    hist_li <- lapply(samples_hist, function(i) x_orig[(i - n_start + 1):i, , drop = FALSE])
  } else {
    if (!is.list(hist_li)) {
      stop("hist_li should be a list of vectors/matrices")
    }
    if (unique(sapply(hist_li, nrow_length)) != n_start) {
      stop("each element of hist_li should have length lags (+1 if VECM)")
    }
  }
  
  samples_hist # index for history
  hist_li # drawn history (of observations)
  
  if (is.null(shock_li)) {
    if (!is.null(seed)) {
      set.seed(seed)
    }
    samples_shock <- sample(seq_len(n_used), size = n.shock, replace = FALSE)
    shock_li <- lapply(samples_shock, function(i) resids[i, , drop = FALSE])
  } else {
    if (!is.list(shock_li)) {
      stop("shock_li should be a list of vectors")
    }
    if (unique(sapply(shock_li, nrow_length)) != 1) {
      stop("each element of shock_li should have length lags")
    }
  }
  
  samples_shock # index for shocks
  shock_li # drawn shocks (of residuals)
  
  M <- expand.grid(hist = hist_li, shock = shock_li)
  n_cases <- nrow(M)
  shock_M <- as.data.frame(do.call("rbind", M$shock))
  colnames(shock_M) <- paste("shock_var", seq_len(nVar), sep = "")
  hist_M <- as.data.frame(do.call("rbind", lapply(M$hist, function(x) c(as.matrix(x)))))
  colnames(hist_M) <- paste("hist_x", rep(seq_len(nVar), each = n_start), "_l", rep(seq_len(n_start), times = nVar), sep = "")
  rep_info <- cbind(n_simu = seq_len(nrow(M)), hist_M, shock_M)
  
  # sims <- lapply(1:nrow(M), function(i) irf_1_shock_ave(object = object, shock = M$shock[[i]], hist = M$hist[[i]], n.ahead = n.ahead, R = R, seed = seed, ...))
  sims <- lapply(1:nrow(M), function(i) tsDyn:::irf_1_shock_ave(object = object, shock = M$shock[[i]], hist = M$hist[[i]], n.ahead = n.ahead, R = R, seed = seed))
  
  sims[[1]]
  sims[[400]]
  
  sims_df <- do.call("rbind", sims)
  n.ahead_here <- length(unique(head(sims_df$n.ahead, 2 * (n.ahead * lag))))
  sims_df$n_simu <- rep(seq_len(n_cases), each = n.ahead_here * nVar)
  sims_df$girf <- with(sims_df, sim_1 - sim_2)
  
  sims_df2 <- merge(rep_info, sims_df, by = "n_simu", sort = FALSE)
  class(sims_df2) <- c("GIRF_df", "data.frame")
  
  # sims_df2
  head(sims_df2, 30)
  tail(sims_df2, 30)
  
# }

# ANCKER
  
# tsDyn_GIRF_setar_fun no 02 ----

# tsDyn_GIRF_setar_fun - function (object, n.ahead = 10, seed = NULL, n.hist = 20, n.shock = 20, 
#                                  R = 10, hist_li = NULL, shock_li = NULL, ...) {
  
  # inputs
  object <- vecm.tsdyn
  n.ahead <- 10
  seed <- NULL
  n.hist = 20
  n.shock = 20
  R = 10
  hist_li <- NULL
  shock_li <- NULL
  
    
  
  # lag <- tsDyn:::get_lag(object)
  lag <- tsDyn:::get_lag.nlVar(object)
  n_start <- ifelse(inherits(object, c("VECM", "TVECM")), lag + 1, lag)
  # x_orig <- get_data_orig(object, as.df = TRUE)
  x_orig <- tsDyn:::get_data_orig.nlVar(object, as.df = TRUE)
  N <- nrow(x_orig)
  resids <- as.data.frame(residuals(object, initVal = FALSE))
  n_used <- nrow(resids)
  # nVar <- get_nVar(object)
  nVar <- tsDyn:::get_nVar.nlVar(object)
  sample_hist <- function() {
    hist_M <- sample(n_start:N, size = 1, replace = FALSE)
    (hist_M - n_start + 1):hist_M
  }
  nrow_length <- function(x) {
    nr <- nrow(x)
    if (is.null(nr)) {
      nr <- length(x)
    }
    nr
  }
  if (is.null(hist_li)) {
    if (!is.null(seed)) {
      set.seed(seed)
    }
    samples_hist <- sample(n_start:N, size = n.hist, replace = FALSE)
    hist_li <- lapply(samples_hist, function(i) x_orig[(i - n_start + 1):i, , drop = FALSE])
  } else {
    if (!is.list(hist_li)) {
      stop("hist_li should be a list of vectors/matrices")
    }
    if (unique(sapply(hist_li, nrow_length)) != n_start) {
      stop("each element of hist_li should have length lags (+1 if VECM)")
    }
  }
  if (is.null(shock_li)) {
    if (!is.null(seed)) {
      set.seed(seed)
    }
    samples_shock <- sample(seq_len(n_used), size = n.shock, replace = FALSE)
    shock_li <- lapply(samples_shock, function(i) resids[i, , drop = FALSE])
  } else {
    if (!is.list(shock_li)) {
      stop("shock_li should be a list of vectors")
    }
    if (unique(sapply(shock_li, nrow_length)) != 1) {
      stop("each element of shock_li should have length lags")
    }
  }
  M <- expand.grid(hist = hist_li, shock = shock_li)
  n_cases <- nrow(M)
  shock_M <- as.data.frame(do.call("rbind", M$shock))
  colnames(shock_M) <- paste("shock_var", seq_len(nVar), sep = "")
  hist_M <- as.data.frame(do.call("rbind", lapply(M$hist, function(x) c(as.matrix(x)))))
  colnames(hist_M) <- paste("hist_x", rep(seq_len(nVar), each = n_start), "_l", rep(seq_len(n_start), times = nVar), sep = "")
  rep_info <- cbind(n_simu = seq_len(nrow(M)), hist_M, shock_M)
  # sims <- lapply(1:nrow(M), function(i) irf_1_shock_ave(object = object, shock = M$shock[[i]], hist = M$hist[[i]], n.ahead = n.ahead, R = R, seed = seed, ...))
  sims <- lapply(1:nrow(M), function(i) tsDyn:::irf_1_shock_ave(object = object, shock = M$shock[[i]], hist = M$hist[[i]], n.ahead = n.ahead, R = R, seed = seed))
  sims_df <- do.call("rbind", sims)
  n.ahead_here <- length(unique(head(sims_df$n.ahead, 2 * (n.ahead * lag))))
  sims_df$n_simu <- rep(seq_len(n_cases), each = n.ahead_here * nVar)
  sims_df$girf <- with(sims_df, sim_1 - sim_2)
  sims_df2 <- merge(rep_info, sims_df, by = "n_simu", sort = FALSE)
  class(sims_df2) <- c("GIRF_df", "data.frame")
  
  # sims_df2
  head(sims_df2, 30)
  tail(sims_df2, 30)
  
# }
  
```

Definition generalized impulse response functions (GIRs) based on Pesaran (2015)

General

\begin{align*}
GI_{y}\left(n, u_{t}^{0}, \Omega_{t-1}^{0}\right) = \text{E}\left(y_{t+n}|u_{t}=u_{t}^{0}, \Omega_{t-1}^{0}\right) - \text{E}\left(y_{t+n}|\Omega_{t-1}^{0}\right)
\end{align*}

Particular VAR

\begin{align*}
GI_{y}\left(n, u_{t}^{0}, \Omega_{t-1}^{0}\right) = A_{n}u_{t}^{0}
\end{align*}

which is independent of the "history" of the process.

The choice of the vector shocks $u_{t}^{0}$ is arbitrary.

One possibility is to consider a large number of likely shocks and then examine the empirical distribution function of $A_{n}u_{t}^{0}$ for all these shocks.

In the case where $u_{t}^{0}$ is drawn from the same distribution as $u_{t}$, namely a multivariate normal distribution with zero means and a constant covariance matrix $\Sigma$ we have the analyitical result

\begin{align*}
GI_{y}\left(n, u_{t}^{0}, \Omega_{t-1}^{0}\right) \sim N\left(0, A_{n} \Sigma A_{n}^{'}\right)
\end{align*}

Persistence profiles (general on X)

Lee  and  Pesaran  (1993) refer  to 
H,(n),  seen  as  a  function  of  n, as  the  (unscaled)  ‘persistence  profile’,  character- 
izing the  time  profile  of the  effects  of the  system-wide  shocks  on  X,.

Persitence profiles (specific on Z)

The  above  concept  of  the  persistence  profile  can  be  readily  extended  to 
cointegrated  systems.  The  (unscaled)  persistence  profile  of  the  cointegrating 
relations  2,  =  P’X,  is given  by 



* `n.hist`:
  * The number of past histories to consider. Should be high, ideally size of data (minus lags).
  * For VAR models ...

Some interpretation:

* 400 simulations *times* 11 steps *times* 2 shock variables *equals* 8800 rows
* `girf = sim_1 - sim_2` 

Check plot methods:

```{r, eval=FALSE}

tsDyn_plot_GIRF_df <- function (x, plot_type = c("density", "line"), n.ahead = c(1, 5, 10), var = unique(x$var)[1], n_simu = c(1, 2), ...) {
  
  # inputs
  x <- sims_df2
  
  plot_type <- "density"
  n.ahead <- c(1, 5, 10)
  var <- unique(x$var)[1]
  n_simu <- c(1, 2)
  
  
  
  # plot_type <- match.arg(plot_type)
  plot_type <- plot_type
  if (plot_type == "density") {
    # plot_GIRF_dens(x = x, n.ahead = n.ahead, var = var, ...)
    tsDyn:::plot_GIRF_dens(x = x, n.ahead = n.ahead, var = var)
  } else if (plot_type == "line") {
    # plot_GIRF_line(x = x, n_simu = n_simu, var = var, ...)
    tsDyn:::plot_GIRF_line(x = x, n_simu = n_simu, var = var)
  }
    
}

tsDyn_plot_GIRF_line <- function (x, n_simu = 1:5, var = unique(x$var)[1], add_legend = TRUE, ...) {
  
  # inputs
  x <- sims_df2
  n_simu <- 1:5
  var <- unique(x$var)[1]
  add_legend <- TRUE
  
  
  
  df <- x[x$n_simu %in% n_simu & x$var %in% var, c("n_simu", "var", "n.ahead", "girf")]
  df_w <- reshape(df, idvar = c("n.ahead", "var"), timevar = "n_simu", direction = "wide", v.names = "girf")
  colnames(df_w) <- gsub("girf\\.", "n_simu: ", colnames(df_w))
  df_w2 <- df_w[, -c(1, 2), drop = FALSE]
  # plot(NA, xlim = range(df$n.ahead), ylim = range(df$girf), ylab = "GIRF", xlab = "n.ahead", ...)
  plot(NA, xlim = range(df$n.ahead), ylim = range(df$girf), ylab = "GIRF", xlab = "n.ahead")
  invisible <- lapply(1:ncol(df_w2), function(x) lines(x = df_w$n.ahead, y = df_w2[, x], col = x))
  if (add_legend) {
    legend("topright", legend = colnames(df_w2), fill = 1:nrow(df_w2))
  }
        
}

tsDyn_plot_GIRF_dens <- function (x, n.ahead = c(1, 5, 10), var = unique(x$var)[1], add_legend = TRUE, ...) {
  
  # inputs
  x <- sims_df2
  n.ahead <- c(1, 5, 10)
  var <- unique(x$var)[1]
  add_legend <- TRUE
  
  
  
  df <- x[x$n.ahead %in% n.ahead & x$var %in% var, c("n_simu", "var", "n.ahead", "girf")]
  df_w <- reshape(df, idvar = c("n_simu", "var"), timevar = "n.ahead", direction = "wide", v.names = "girf")
  colnames(df_w) <- gsub("girf\\.", "n.ahead: ", colnames(df_w))
  dens <- apply(df_w[, -c(1, 2), drop = FALSE], 2, density)
  # plot(NA, xlim = range(sapply(dens, "[", "x")), 
  #     ylim = range(sapply(dens, "[", "y")), ylab = "GIRF", 
  #     xlab = "Density", ...)
  plot(NA, xlim = range(sapply(dens, "[", "x")), 
      ylim = range(sapply(dens, "[", "y")), ylab = "GIRF", 
      xlab = "Density")
  invisible <- mapply(lines, dens, col = 1:length(dens))
  if (add_legend) {
    legend("topright", legend = names(dens), fill = 1:length(dens))
  }
        
}

```





<hr>

Step x: Call function `tsDyn:::irf_1_shock_ave()`

```{r, cache=TRUE}

# tsDyn_irf_1_shock_ave <- function (object, shock, hist, R = 10, n.ahead = 10, innov = NULL, 
#                                    shock_both = TRUE, returnStarting = FALSE, add.regime = FALSE, 
#                                    seed = NULL) {
  
  # inputs
  i <- 1
  object <- vecm.tsdyn
  shock <- M$shock[[i]]
  hist <- M$hist[[i]]
  n.ahead <- 10
  R <- 10
  seed <- NULL
  
  innov <- NULL
  shock_both <- TRUE
  returnStarting <- FALSE
  add.regime <- FALSE

  
  
  if (!is.null(seed)) {
    set.seed(seed)
  }
  # out <- replicate(R, irf_1_shock(object = object, shock = shock, 
  #                                 hist = hist, n.ahead = n.ahead, innov = innov, shock_both = shock_both, 
  #                                 returnStarting = returnStarting, add.regime = add.regime), 
  #                  simplify = FALSE)
  out <- replicate(R, tsDyn:::irf_1_shock(object = object, shock = shock, 
                                          hist = hist, n.ahead = n.ahead, innov = innov, shock_both = shock_both, 
                                          returnStarting = returnStarting, add.regime = add.regime), 
                   simplify = FALSE)
  out_M <- do.call("rbind", out)
  out_M_means <- aggregate(out_M[, grep("sim|regime", colnames(out_M))], list(n.ahead = out_M$n.ahead, var = out_M$var), mean)
  out_M_means
    
# }

```

Step x: Call function `tsDyn:::irf_1_shock()`

```{r, cache=TRUE}

# tsDyn_irf_1_shock_fun <- function (object, shock, hist,
#                                    n.ahead = 10, innov = NULL, shock_both = TRUE, 
#                                    returnStarting = FALSE, add.regime = FALSE, seed = NULL) {
  
  # inputs
  object <- vecm.tsdyn
  i <- 1
  shock <- M$shock[[i]]
  hist <- M$hist[[i]]
  
  n.ahead <- 10
  innov <- NULL
  shock_both <- TRUE
  returnStarting <- FALSE
  add.regime <- FALSE
  add.regime <- FALSE
  seed <- NULL
  

  
  # lag <- get_lag(object)
  lag <- tsDyn:::get_lag.nlVar(object)
  n_start <- ifelse(inherits(object, c("VECM", "TVECM")), lag + 1, lag)
  include <- object$include
  B <- coef(object, hyperCoef = FALSE)
  # nthresh <- get_nthresh(object)
  nthresh <- tsDyn:::get_nthresh.nlVar(object)
  # Thresh <- getTh(object)
  Thresh <- tsDyn:::getTh.nlVar(object)
  N <- n.ahead + 1
  # K <- get_nVar(object)
  K <- tsDyn:::get_nVar.nlVar(object)
  # series <- get_series(object)
  series <- tsDyn:::get_series(object)
  beta <- object$model.specific$coint
  model <- switch(class(object)[[1]],
                  ar = "setar",
                  linear = "setar", 
                  setar = "setar",
                  VAR = "TVAR",
                  TVAR = "TVAR", 
                  VECM = "TVECM",
                  TVECM = "TVECM", stop("Error model not recognised!"))
  if (inherits(object, c("TVAR", "TVECM"))) {
    B <- do.call("cbind", B)
  }
  if (is.null(innov)) {
    res_obj <- as.matrix(residuals(object))
    res_obj <- res_obj[-seq_len(lag), , drop = FALSE]
    if (!is.null(seed)) {
      set.seed(seed)
    }
    index_samp <- sample(seq_len(nrow(res_obj)), N, replace = FALSE)
    innov <- res_obj <- res_obj[index_samp, , drop = FALSE]
  }
  hist_M <- as.matrix(hist)
  shock_M <- as.matrix(shock)
  if (nrow(hist_M) != n_start) {
    stop("hist should be of same nrow as lag (+1 if VECM)")
  }
  if (ncol(hist_M) != K) {
    stop("hist should be of same ncol as number of variables")
  }
  if (nrow(shock_M) != 1) {
    stop("shock should have only one row")
  }
  if (ncol(shock_M) != K) {
    stop("shock should be of same ncol as number of variables")
  }
  if (nrow(innov) != N) {
    stop("innov should be of same length as n.ahead + 1")
  }
  innov_1 <- rbind(shock_M, innov[-1, , drop = FALSE])
  if (shock_both) {
    innov_1 <- innov + rbind(shock_M, matrix(0, nrow = n.ahead, ncol = K))
  }
  innov_2 <- innov
  # sim_1 <- model.gen(model = model, B = B, lag = lag, include = include, 
  #                    nthresh = nthresh, Thresh = Thresh, beta = beta, starting = hist_M, 
  #                    innov = innov_1, n = N, returnStarting = TRUE, add.regime = add.regime)
  sim_1 <- tsDyn:::model.gen(model = model, B = B, lag = lag, include = include, 
                     nthresh = nthresh, Thresh = Thresh, beta = beta, starting = hist_M, 
                     innov = innov_1, n = N, returnStarting = TRUE, add.regime = add.regime)
  # sim_2 <- model.gen(model = model, B = B, lag = lag, include = include, 
  #                    nthresh = nthresh, Thresh = Thresh, beta = beta, starting = hist_M, 
  #                    innov = innov_2, n = N, returnStarting = TRUE, add.regime = add.regime)
  sim_2 <- tsDyn:::model.gen(model = model, B = B, lag = lag, include = include, 
                     nthresh = nthresh, Thresh = Thresh, beta = beta, starting = hist_M, 
                     innov = innov_2, n = N, returnStarting = TRUE, add.regime = add.regime)
  
  n.aheads_all <- c(-n_start:0, seq_len(n.ahead))
  if (K > 1) {
    df <- data.frame(var = rep(series, each = nrow(sim_1)), n.ahead = n.aheads_all, sim_1 = unlist(sim_1), sim_2 = unlist(sim_2))
    rownames(df) <- seq_len(nrow(df))
  } else {
    df <- data.frame(var = series, n.ahead = n.aheads_all, sim_1 = sim_1$res, sim_2 = sim_2$res)
  }
  if (add.regime) {
    df <- df[, 1:4]
    colnames(df) <- c("n.ahead", "sim_1", "regime_1", "sim_2")
  }
  if (!returnStarting) {
    df <- subset(df, n.ahead >= 0)
  }
  
  df
  
# }

```

```{r}

tsDyn_model_gen_fun <- function (model = c("setar", "TVAR", "TVECM"), B, n = 200, lag = 1, 
                                 include = c("const", "trend", "none", "both"), nthresh = 0, 
                                 thDelay = 0, Thresh, trendStart = 1, starting = NULL, innov, 
                                 exo, round_digits = 10, returnStarting = FALSE, add.regime = FALSE, 
                                 show.parMat = FALSE, thVar = NULL, mTh = 1, beta, ...) {
  
  model <- model
  B <- B
  lag <- lag
  include <- include
  nthresh <- nthresh
  Thresh <- Thresh
  beta <- beta
  starting <- hist_M
  innov <- innov_1
  n <- N
  
  thDelay <- 0
  trendStart <- 1
  exo <- NULL
  round_digits <- 10
  returnStarting <- TRUE
  add.regime <- add.regime
  show.parMat <- FALSE
  thVar <- NULL
  mTh <- 1 
  
  
  
  # model <- match.arg(model)
  model <- model
  if (model == "setar") {
    res <- tsDyn:::setar.gen(B = B, n = n, lag = lag, include = include, 
                             nthresh = nthresh, Thresh = Thresh, thDelay = thDelay, 
                             starting = as.numeric(starting[, 1]), innov = innov, 
                             exo = exo, trendStart = trendStart, round_digits = round_digits, 
                             returnStarting = returnStarting, add.regime = add.regime, 
                             show.parMat = show.parMat, ...)
  } else if (model == "TVAR") {
    res <- tsDyn:::TVAR.gen(B = B, n = n, lag = lag, include = include, 
                            nthresh = nthresh, Thresh = Thresh, thDelay = thDelay, 
                            starting = starting, innov = innov, thVar = thVar, 
                            mTh = mTh, trendStart = trendStart, round_digits = round_digits, 
                            returnStarting = returnStarting, add.regime = add.regime, 
                            show.parMat = show.parMat, ...)
  } else if (model == "TVECM") {
    res <- tsDyn:::TVECM.gen(B = B, n = n, lag = lag, include = include, 
                             beta = beta, nthresh = nthresh, Thresh = Thresh, 
                             starting = starting, innov = innov, round_digits = round_digits, 
                             returnStarting = returnStarting, show.parMat = show.parMat, ...)
  }
  as.data.frame(res)
}

```

```{r}

tsDyn_TVECM_gen_fun <- function (B, beta, n = 200, lag = 1,
                                 include = c("const", "trend", "none", "both"), nthresh = 1, Thresh,
                                 starting = NULL, innov, returnStarting = FALSE, show.parMat = FALSE, round_digits = 4) {
  
  # inputs
  B <- B
  n <- n
  lag <- lag
  include <- include
  beta <- beta
  nthresh <- nthresh
  Thresh <- Thresh
  starting <- starting
  innov <- innov
  round_digits <- round_digits
  returnStarting <- returnStarting
  show.parMat <- show.parMat
  
  

  p <- lag
  # include <- match.arg(include)
  include <- include
  if (!nthresh %in% c(0, 1, 2)) {
    stop("Arg nthresh should  be either 0, 1 or 2")
  }
  if (missing(beta)) {
    stop("please provide arg beta (cointegrating value)")
  }
  ninc <- switch(include, none = 0, const = 1, trend = 1, both = 2)
  incVal <- switch(include, none = NULL, const = "const", trend = "trend", both = c("const", "trend"))
  k <- nrow(B)
  if (is.vector(beta)) {
    if (length(beta) == k - 1) {
      beta <- c(1, -beta)
    }
    tBETA <- matrix(beta, nrow = 1)
    r <- 1
  } else {
    if (nrow(beta) != k) {
      stop("beta should have k rows and r cols")
    }
    r <- ncol(beta)
    tBETA <- t(beta)
  }
  esp <- p * k + r + ninc
  pa <- switch(as.character(nthresh), `0` = "", `1` = c("_low", "_upr"), `2` = c("_low", "_mid", "_upr"))
  lags <- as.vector(outer("L{x", 1:k, paste, sep = ""))
  lags2 <- paste(rep(lags, times = p), "}{", rep(1:p, each = p), "}", sep = "")
  if (esp * (nthresh + 1) != ncol(B)) {
    colnames_Matrix_input <- as.vector(outer(c(rep("ECT", r), incVal, lags2), pa, paste, sep = ""))
    cat("Matrix B badly specified: should have ", esp * (nthresh + 1), "columns, but has", ncol(B), "\n")
    print(matrix(NA, nrow = k, ncol = esp * (nthresh + 1), dimnames = list(paste("Equ x", 1:k, sep = ""), colnames_Matrix_input)))
    stop()
  }
  rownames(B) <- paste("Equ x", 1:k, ":", sep = "")
  if (!is.null(starting)) {
    if (!is.matrix(starting)) {
      stop("Provide 'starting' as matrix")
    }
    if (!all(dim(starting) == c(p + 1, k))) {
      stop("Bad specification of starting values. Should have nrow = lag +1 and ncol = number of variables")
    }
  }
  Bmat <- B
  if (include != "both") {
    aa1 <- r + switch(include, none = 1:2, const = 2, trend = 1, both = NULL)
    aa <- sort(rep(aa1, each = nthresh + 1) + (0:nthresh) * (p * k + max(aa1)))
    # Bmat <- myInsertCol(Bmat, c = aa, 0)
    Bmat <- tsDyn:::myInsertCol(Bmat, c = aa, 0)
  }
  nparBmat <- p * k + 2 + r
  Yb <- matrix(0, nrow = n + p + 1, ncol = k)
  if (!is.null(starting)) {
    Yb[1:(p + 1), ] <- starting
  }
  trend <- c(rep(NA, p + 1), seq_len(n))
  
  # start generation of Yb
  
  if (nthresh == 0) {
    for (i in (p + 2):(n + p + 1)) {
      ECT <- Bmat[, 1:r] %*% tBETA %*% matrix(Yb[i - 1, ], ncol = 1)
      Yb[i, ] <- rowSums(cbind(Yb[i - 1, ], Bmat[, r + 1], Bmat[, r + 2] * trend[i], ECT, Bmat[, -c(1:(r + 2))] %*% matrix(t(Yb[i - c(1:p), ] - Yb[i - c(2:(p + 1)), ]), ncol = 1), innov[i - p - 1, ]))
    }
  } else if (nthresh == 1) {
    BD <- Bmat[, seq_len(nparBmat)]
    BU <- Bmat[, -seq_len(nparBmat)]
    for (i in (p + 2):(n + p + 1)) {
      ECT <- Yb[i - 1, , drop = FALSE] %*% t(tBETA)
      if (round(ECT, round_digits) <= Thresh) {
        B_here <- BD
      } else {
        B_here <- BU
      }
      Yb[i, ] <- rowSums(cbind(Yb[i - 1, ], B_here[, 1] * as.vector(ECT), B_here[, 2], B_here[, 3] * trend[i], B_here[, -c(1, 2, 3)] %*% matrix(t(Yb[i - c(1:p), ] - Yb[i - c(2:(p + 1)), ]), ncol = 1), innov[i - p - 1, ]))
    }
  } else if (nthresh == 2) {
    BD <- Bmat[, seq_len(nparBmat)]
    BM <- Bmat[, seq_len(nparBmat) + nparBmat]
    BU <- Bmat[, seq_len(nparBmat) + 2 * nparBmat]
    for (i in (p + 2):(n + p + 1)) {
      ECT <- Yb[i - 1, , drop = FALSE] %*% t(tBETA)
      if (round(ECT, round_digits) <= Thresh[1]) {
        B_here <- BD
      } else if (round(ECT, round_digits) > Thresh[2]) {
        B_here <- BU
      } else {
        B_here <- BM
      }
      Yb[i, ] <- rowSums(cbind(Yb[i - 1, ], B_here[, 1] * as.vector(ECT), B_here[, 2], B_here[, 3] * trend[i], B_here[, -c(1, 2, 3)] %*% matrix(t(Yb[i - c(1:p), ] - Yb[i - c(2:(p + 1)), ]), ncol = 1), innov[i - p - 1, ]))
    }
  }
  if (show.parMat) {
    colnames_Matrix_system <- as.vector(outer(c(rep("ECT", r), "Const", "Trend", lags2), pa, paste, sep = ""))
    colnames(Bmat) <- colnames_Matrix_system
    print(Bmat)
  }
  if (!returnStarting) {
    Yb <- Yb[-seq_len(p + 1), ]
  }
      
  return(Yb)
  
}

```

Strategies to hijack `tsDyn:::GIRF` function

* edit `tsDyn::TVECM` function for known breaks to create ingredients of `tsDyn:::GIRF` function.



## TVECM model estimation

```{r}

tvec <- tsDyn::TVECM(tsDyn::zeroyldMeta[, c("long.run", "short.run")], lag = 2, ngridTh = 300, beta = list(exact = 1))

```

## TVECM model orthngalized IRF analysis

```{r, eval=FALSE}

# don't run
irf.tvec <- tsDyn:::irf.VECM(tvec, ortho = FALSE)

```

```{r}

irf.tvec <- tsDyn:::irf.TVECM(tvec, ortho = TRUE, boot = FALSE, n.ahead = 10)
irf.tvec$irf$short.run
irf.tvec$irf$long.run

irf.l.tvec <- tsDyn:::irf.TVECM(tvec, ortho = TRUE, regime = "L", boot = FALSE, n.ahead = 10)
irf.l.tvec$irf$short.run
irf.l.tvec$irf$long.run

irf.h.tvec <- tsDyn:::irf.TVECM(tvec, ortho = TRUE, regime = "H", boot = FALSE, n.ahead = 10)
irf.h.tvec$irf$short.run
irf.h.tvec$irf$long.run

```

## TVECM model generalized IRF analysis


